---
title: "Quantifying the propensity of forest reorganization"
author: "Mária Potterf"
date: "8/18/2022"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, tidy = TRUE, tidy.opts = list(comment = FALSE), message = F)
knitr::opts_chunk$set(fig.width=6, fig.height=2.5) 

```

# Background and implementation:

identify the aspects of the forest reorganization:

- in forest composition
- in forest structure

evaluate to understand if there are changes towards one of the axes:

- reassembly
- restructure
- replacement
- resilience   


# Describe study sites: 

The number of samples triplets by dominant tree species. Each triplet has 3 management categories: 

- Cleared
- Dead
- Living


```{r read-libs, eval = TRUE, echo=FALSE, include = FALSE, warning = FALSE, error = FALSE }
# rm(list=ls())

# Input data -------------------------------------------------------------------
getwd()
load(file = paste(getwd(), "dataToPlot.Rdata", sep = '/'))


#### Source paths and functions  -----------------------------------------------

source('myPaths.R')


#### Read libraries  -----------------------------------------------------------
library(knitr)
library(readxl)
library(dplyr)
library(data.table)
library(tidyr)
library(ggplot2)
library(stringr)  # use regex expressions
#library(gridExtra)
library(ggpubr)

##### Stats
library(MuMIn)
library(vegan)
library(mgcv)
library(gratia) # visualization of mcv


## Colors
cols = c('#0072B2', # 
         '#E69F00', # 
         '#F0E442', # 
         '#000000') # black: spruce


```


```{r triplet-count, echo=FALSE}
out_summary <- dat %>% 
  group_by(dom_sp) %>% 
  distinct(trip_n) %>% 
  tally() %>% 
  rename(species=dom_sp,
         triplets_count = n)

```


```{r sites-count, echo=FALSE}
n_sites <- 
  dat %>% 
  group_by(trip_n, dom_sp, manag, sub_n) %>% 
  nrow() 

```

**Sites overview:** 

Triplets count:  **`r length(unique(dat$trip_n))`**

Total count of plots (4m2) per triplet&category : **`r n_sites`** plots



```{r print-summary, echo=F, caption = 'salala' }
# print nice table in knitr
out_summary %>% knitr::kable("markdown")

```

### Prevalence of species in regeneration by height category:

Tree regeneration heights classes:

- seedlings: 0.2-2m
- saplings: 2-5m, up to 10 cm dbh

Tree regeneration notes: 

- O_soft = other softwood species
- O_hard = other hardwood species

```{r p_prevalence, echo = F, fig.cap = 'Fig. Average counts per tree species. Ordered by the most aboundant tree species'}
p_prevalence

```



#### Species composition by dominant tree species and category (for seedlings, saplings)

```{r plot-regen-sp-categ, echo = F, fig.height=5}
p_treeComp_categ

```


### get importance value only for the regeneration?

#### Novel species: Presence & absence
Identify the tree species that are novel (appear after the disturbance) or absent (disappear) after disturbances. Compared across triplets between reference (living) and disturbed conditions.

```{r p-winners-loosers, fig.height=4}
p_win_loos_sp

```





#### Novel species: competition

