---
title: "Sites overview Franconia"
author: "Maria Potterf"
date: "7/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, tidy = TRUE, tidy.opts = list(comment = FALSE), message = F)
knitr::opts_chunk$set(fig.width=6, fig.height=2.5) 
```

#### MS: Early indication of post-disturbance reorganization in Central European forests

The number of samples triplets by dominant tree species. Each triplet has 3 management categories: 

- Cleared
- Dead
- Living

```{r read-libs, eval = TRUE, echo=FALSE, include = FALSE, warning = FALSE, error = FALSE }
rm(list=ls())



# Input data -------------------------------------------------------------------

load(file = "vegData.Rdata")


#### Source paths and functions  -----------------------------------------------

source('myPaths.R')


#### Read libraries  -----------------------------------------------------------
library(knitr)
library(readxl)
library(dplyr)
library(data.table)
library(tidyr)
library(ggplot2)
library(stringr)  # use regex expressions
#library(gridExtra)
library(ggpubr)

##### Stats
library(MuMIn)
library(vegan)
library(mgcv)
library(gratia) # visualization of mcv


## Colors
cols = c('#0072B2', # strict reserves
         '#E69F00', # buffer 500
         '#F0E442', # buffer 2000
         '#000000') # control


```



### Triplets counts

Currently we have 20 triplets samples, from total of 40. Stratification by dominant tree species:

```{r triplet-count, echo=FALSE}
out_summary <- dat %>% 
  group_by(dom_sp) %>% 
  distinct(trip_n) %>% 
  tally() %>% 
  rename(species=dom_sp,
         triplets_count = n)

```

```{r print-summary, echo=F}
# print nice table in knitr
out_summary %>% knitr::kable("markdown")

```



####  Variables:

- forest structure: 

  - density of regeneration (here less< 2m height)
  - height diversity
  
- forest composition: 

  - Shannon index
  - effective number of species
  
- ecological traits (based on Niinements (2006)), community weighted means (CWM) 

  - CWM shading tolerance
  - CWM drought tolerance



### Regeneration density 
```{r slope-corr, include = F}

# Example of slope correction calculation of tree density:
ha = 10000
trees_field = 10
gradient = 16.7  #(has to be in degrees!)

# area of the subsite: 2x2 m
r = 2    # m
r1 = r
r2 = r1*cos(gradient*pi/180)   # R works in radians: to get the value in degrees, it has to be in form cos(angle * pi/180) 
# https://r-lang.com/r-cos-function-with-example/
area_field = r^2# m2
area_plane = r1*r2

# Get the correction factor:
ideal_factor   = ha/area_field
correct_factor = ha/area_plane

# Calculate teh tree deisnity based on field, and based on corrected area:
trees_dens_field = trees_field*ideal_factor
trees_dens_plane = trees_field*correct_factor


```


```{r get-calc-dens, include = F}

# Define sample area per patch - correct the density/ha estimation----------------
# calculate from the original data table
subsample_n <- 
  dat %>%
  group_by(trip_n , dom_sp , manag ) %>% 
  distinct(sub_n ) %>% 
  tally() %>%
  mutate(trip_n = as.character(trip_n)) %>% 
  rename(n_subsites = n)



# The correction of the counts: need to do it on the level of individual subsites:
# as there is high variation in slopes ('gradient') between the subsites
# So I am correcting the tree density/ha per each subsite: 
# maybe then use teh mean/sum densities per category??? 
df_regen <- df_regen %>% 
  mutate(manag = factor(manag, levels = c('l', 'c', 'd'),
                        labels = c('living','cleared', 'dead'))) %>% 
  mutate(dom_sp = factor(dom_sp))



# Get tree density/ha across all heights, correct the density by slope per each subsite
df_reg_dens <- 
  df_regen %>% 
  left_join(subsample_n) %>% 
  ungroup() %>% 
  group_by(trip_n, dom_sp, manag, n_subsites, height_class, reg_species)  %>% 
    mutate(length_corr = 2*cos(gradient*pi/180),
           area_corr   = 2*length_corr,
           correct_factor = ha/area_corr,
           corr_density = n_total*correct_factor)  %>% 
    filter(corr_density != 0) %>% 
  ungroup(.) #%>% 


```



```{r, calc-summary-stat, include = F}
# get summary statictics: 
# how much regeneration is per each site? now it is splitted among several species on site
# but it is calculated for each species as a value per hectar. So the means should be good here (accounting for species numbers)
df_out <- df_reg_dens %>% 
  group_by(manag, dom_sp) %>% 
  summarize(mean_dens = mean(corr_density, na.rm = T),
            sd_dens   = sd(corr_density, na.rm = T)) %>% 
  mutate(dens_mean_sd = stringr::str_glue("{round(mean_dens,1)}±{round(sd_dens,1)}")) %>% 
  dplyr::select(manag, dom_sp, dens_mean_sd) %>% 
  pivot_wider(names_from = manag, 
              values_from = dens_mean_sd )

```  

Regeneration density of the forest regeneration (<2m). Average values per subsites.

```{r print-summary-dens, echo = F, include = T}
df_out %>% knitr::kable("markdown")


```

```{r plot-density, echo = F}
df_reg_dens %>%
  ggplot(aes(x = dom_sp, 
             y = corr_density/10000)) +
  theme_bw() +
  stat_summary(geom = "errorbar",
               width = 0.5,
               aes(col = dom_sp)) +
  stat_summary(geom = "pointrange", aes(col = dom_sp)) +
  scale_colour_manual(values=cols, 
                      name="Dominant\nspecies") + 
  facet_grid(~manag) +
  labs(y='Regeneration density*10000')

```
  



#### Shannon index: effective number of species

```{r shannon-calc, echo = F, include = F, warning=FALSE}
df_reg_dens_shannon <-
  df_reg_dens %>%
  group_by(trip_n, dom_sp, manag, reg_species) %>%
  summarize(mean_dens = mean(corr_density)) %>% # means first I have values for individual species and subsites!
  mutate(
    dens_tot = sum(mean_dens),
    sp_pi = mean_dens / dens_tot,
    shannon_part = sp_pi * log(sp_pi)
  ) %>%
  summarize(shannon = -sum(shannon_part)) %>%
  mutate(effective_n_sp = exp(shannon)) #%>%
  
```




```{r plot-shannon, echo = F, include = T}


df_reg_dens_shannon %>% 
  ggplot(aes(y = effective_n_sp,
             x = dom_sp)) +
#  geom_boxplot() + 
  theme_bw() +
  stat_summary(geom = "errorbar",
               width=0.5,
               aes(col = dom_sp)) +
  stat_summary(geom = "pointrange", 
               aes(col = dom_sp)) +
  scale_colour_manual(values=cols, 
                      name="Dominant\nspecies") + 
  
  facet_grid(~manag, scales = 'free') +
  ylab('Effective #\nof species')
```





### Community traits

**shade tolerance**: 

  - higher number = more tolerance (fir), 
  - lower number = less tolerence (more sunny site, pine)
  
**drought tolerance**: 
 
  - higher = more tolerance (pine), 
  - lower  = less (more drought sensitive, spruce)


```{r get-traits, include = F}

# Get tree species traits:  ----------------------------------------------------------

eco_traits <- read_excel(paste(myPath,
                               'notes/litterature/traits_database',  
                               'Niinemets_2006.xls', sep = '/'),
                         skip = 3,
                         sheet = 'Niinemets_2006_appendix',
                         .name_repair = function(x) gsub("\\s+", "_", x)) # replace the spaces in colnames by '_'




# Interpretation: 
# shade tolerance: 
#             higher number = more tolerance (fir), 
#             lower number = less tolarence (more sunny site, pine)
# drought tolerance: 
#             higher  = more tolerance (pine), 
#             lower = less (more drought sensitive, spruce)

# Filter only relevant species:  # how to handle the Other sftwoos and Other harvwood? now just skipped 
# Any way it is likely not dominant
trees_lat <- c(
  'Picea abies',
  'Fagus sylvatica',
  'Sorbus aucuparia',
  'Abies alba',
  'Acer pseudoplatanus',
  'Betula pendula',
  'Pinus sylvestris',
  'Fraxinus excelsior'
) 

quercus_spp <- c('Quercus petraea',  # Quercus will be averaged later
                 'Quercus robur')

salix_spp <- c('Salix caprea',     # Salix will be averaged later
               'Salix alba')

# Get Quercus spp: average the values for the :
# Quercus petraea , Quercus robur
# For salix: Salix alba, Salix caprea  (S. fragilis is not that common)

#'OtherHardwood',
# 'OtherSoftwood'

# Filter eco traits database by species: 
# need to do individuall for Quercis, salix and for other species 
# that we have full name identification
traits_Qc <- 
  eco_traits %>% 
  dplyr::select(c('Species','Shade_tolerance', 'Drought_tolerance')) %>% 
  filter(Species %in% quercus_spp)  %>% 
    mutate(Species = 'Quercus') %>% 
    group_by(Species) %>% 
    summarize(Shade_tolerance = mean(Shade_tolerance),
              Drought_tolerance = mean(Drought_tolerance))

traits_Sx <- 
  eco_traits %>% 
  dplyr::select(c('Species','Shade_tolerance', 'Drought_tolerance')) %>% 
  filter(Species %in% salix_spp)  %>% 
  mutate(Species = 'Salix') %>% 
  group_by(Species) %>% 
  summarize(Shade_tolerance   = mean(Shade_tolerance),
            Drought_tolerance = mean(Drought_tolerance))

# remianing species:
traits_sp <- 
  eco_traits %>% 
  dplyr::select(c('Species','Shade_tolerance', 'Drought_tolerance')) %>% 
  filter(Species %in% trees_lat)

# Merge traits into single df
trait_df <- rbind(traits_Qc,
                  traits_Sx,
                  traits_sp) 

# Change naming to be able to merge them with denity dataset:
trait_df <- trait_df %>%
  mutate(
    Species = case_when(
      Species == "Fraxinus excelsior"    ~ "Ash",
      #Species == "Sonstiges NH"         ~ "OtherSoftwood",
      # Species == "Sonstiges LH"        ~ "OtherHardwood",
      Species == "Fagus sylvatica"       ~ "Beech" ,
      Species == "Sorbus aucuparia"      ~ "Rowan",
      Species == "Acer pseudoplatanus"   ~ "Maple",
      Species == "Picea abies"           ~ "Spruce",
      Species == "Quercus"               ~ "Oak",
      Species == "Pinus sylvestris"      ~ "Pine",
      Species == "Betula pendula"        ~ "Birch",
      Species == "Salix"                 ~ "Willow",
      Species == "Abies alba"            ~ "Fir"
    )
  ) %>% 
  rename(reg_species = Species)

```

```{r plot-traits}
# Community weighted means ---------------------------------------------
# https://rpubs.com/CPEL/cwm

df_traits_cwm <- 
  df_reg_dens %>%
  group_by(trip_n, dom_sp, manag, reg_species) %>%
  summarize(mean_dens = mean(corr_density)) %>%
  left_join(trait_df) %>% 
  ungroup(.) %>% 
  group_by(trip_n, dom_sp, manag) %>% 
  summarize(shade_cwm = weighted.mean(Shade_tolerance,     mean_dens, na.rm = TRUE  ),
            drought_cwm = weighted.mean(Drought_tolerance, mean_dens, na.rm = TRUE  )) %>% 
  mutate(manag = factor(manag, 
                        #levels = c('l', 'c', 'd'),
                        levels = c('living','cleared', 'dead')))

# Make some plots:
# set desired dodge width - to keep the error lines apart
pd <- position_dodge(width = 0.4)

p_shade <- 
  df_traits_cwm %>%
  ggplot(aes(x = manag,
             y = shade_cwm))  +
  theme_bw() +
  stat_summary(#geom = "mean", 
    geom = "line", 
    size=0.5, 
    aes(colour = dom_sp, 
        group = dom_sp,
        lty = dom_sp),
    position = pd) +
  stat_summary(geom = "errorbar", 
               width = 0.3,
               aes(col = dom_sp), 
               position = pd) +
  stat_summary(geom = "pointrange", 
               aes(col = dom_sp),
               position = pd) +
  scale_colour_manual(values=cols, 
                      name="Dominant\nspecies") + 
  #facet_grid(~manag) +
  labs(y='Shade tolerance')

# 

p_drought <- 
  df_traits_cwm %>%
  ggplot(aes(x = manag,
             y = drought_cwm))  +
  theme_bw() +
  stat_summary(#geom = "mean", 
               geom = "line", 
               size=0.5, 
               aes(colour = dom_sp, 
                   group = dom_sp,
                   lty = dom_sp),
               position = pd) +
  stat_summary(geom = "errorbar", 
               width = 0.3,
               aes(col = dom_sp), 
               position = pd) +
  stat_summary(geom = "pointrange", 
               aes(col = dom_sp),
               position = pd) +
  scale_colour_manual(values=cols, 
                      name="Dominant\nspecies") + 
  labs(y='Drought tolerance')




ggarrange(p_shade,p_drought, ncol = 2, nrow = 1 , 
          common.legend = TRUE  )



```
#### Changing status:

Reorganization:

- in structure
- in composition

Outcomes:

- no change -> resilience
- in structure -> restructure
- in composition -> reassemble
- in structure&composition -> replace/shift

