---
title: "Sites overview Franconia"
author: "Maria Potterf"
date: "7/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, tidy = TRUE, tidy.opts = list(comment = FALSE), message = F)
knitr::opts_chunk$set(fig.width=6, fig.height=2.5) 
```

#### MS: Early indication of post-disturbance reorganization in Central European forests

The number of samples triplets by dominant tree species. Each triplet has 3 management categories: 

- Cleared
- Dead
- Living

```{r read-libs, eval = TRUE, echo=FALSE, include = FALSE, warning = FALSE, error = FALSE }
# rm(list=ls())

# Input data -------------------------------------------------------------------
getwd()
load(file = paste(getwd(), "dataToPlot.Rdata", sep = '/'))


#### Source paths and functions  -----------------------------------------------

source('myPaths.R')


#### Read libraries  -----------------------------------------------------------
library(knitr)
library(readxl)
library(dplyr)
library(data.table)
library(tidyr)
library(ggplot2)
library(stringr)  # use regex expressions
#library(gridExtra)
library(ggpubr)

##### Stats
library(MuMIn)
library(vegan)
library(mgcv)
library(gratia) # visualization of mcv


## Colors
cols = c('#0072B2', # 
         '#E69F00', # 
         '#F0E442', # 
         '#000000') # black: spruce


```



### Triplets counts

Currently we have 42 triplets. Stratification by dominant tree species:

```{r triplet-count, echo=FALSE}
out_summary <- dat %>% 
  group_by(dom_sp) %>% 
  distinct(trip_n) %>% 
  tally() %>% 
  rename(species=dom_sp,
         triplets_count = n)

```



```{r print-summary, echo=F}
# print nice table in knitr
out_summary %>% knitr::kable("markdown")

```

#### Number of sites per triplet/category: 

```{r sites-count, echo=FALSE}
n_sites <- 
  dat %>% 
  group_by(trip_n, dom_sp, manag, sub_n) %>% 
  nrow() #%>% 
  #pull()
  #distinct() #%>% 
  #tally() #%>% 
  #rename(species=dom_sp,
  #       triplets_count = n)

```

In total, we have `n_sites` sites (~ 10 sites per triplet& category).


#### Species composition



####  Variables:

- forest structure: 
  - density of regeneration (here less< 2m height)
  - height diversity*to be completed (not in report)

- forest composition: 

  - Shannon index 
  - effective number of species 
  
- ecological traits (based on Niinements (2006)), community weighted means (CWM) 

  - CWM shading tolerance
  - CWM drought tolerance



### Regeneration density 


```{r, calc-summary-stat, include = F}
# get summary statictics: 
# how much regeneration is per each site? now it is splitted among several species on site
# but it is calculated for each species as a value per hectar. So the means should be good here (accounting for species numbers)
df_out <- df_reg_dens %>% 
  group_by(manag, dom_sp) %>% 
  summarize(mean_dens = mean(corr_density, na.rm = T),
            sd_dens   = sd(corr_density, na.rm = T)) %>% 
  mutate(dens_mean_sd = stringr::str_glue("{round(mean_dens,1)}±{round(sd_dens,1)}")) %>% 
  dplyr::select(manag, dom_sp, dens_mean_sd) %>% 
  pivot_wider(names_from = manag, 
              values_from = dens_mean_sd )

```  

Regeneration density of the forest regeneration (<2m). Each subsite have been corrected for the differences in slope. The sum values of regenerated tree species counts were divided by the number of subsites. 

```{r print-summary-dens, echo = F, include = T}
df_out %>% knitr::kable("markdown")


```

```{r plot-density, fig.height = 5}

p_density

```


Regeneration density (number of stems per site for trees > 20 cm and <2m tall). All tre species are grouped. Plots represent means +- standard error. 

```{r plot-reg-density, echo = F}
p_reg_dens

```
  



#### Shannon index: 

Plots represent means +- standard error. 


```{r plot-shannon, echo = F, include = T}

p_shannon

```

#### Shannon index: effective number of species

Plots represent means +- standard error. 

```{r plot-eff_number, echo = F, include = T}

p_eff_number
```




### Community traits

**shade tolerance**: 

  - higher number = more tolerance (fir), 
  - lower number = less tolerence (more sunny site, pine)
  
**drought tolerance**: 
 
  - higher = more tolerance (pine), 
  - lower  = less (more drought sensitive, spruce)




```{r plot-traits}
# Community weighted means ---------------------------------------------
# https://rpubs.com/CPEL/cwm





ggarrange(p_shade,p_drought, ncol = 2, nrow = 1 , 
          common.legend = TRUE  )




```

#### Reorganization:

- in structure* to be completed
- in composition

Outcomes for the paper:

- no change -> resilience
- in structure -> restructure
- in composition -> reassemble
- in structure&composition -> replace/shift

Outcomes here just indicates if the regeneration species correspond to dominant species in the matre tree composition. 

If the density of the regeneration tree species:

 - more than 50% of stems: **resilience**
 - if 25-50%: **decrease**
 - if 0-25%: **species change** -> changes in compositions

How do the sites change their composition?

```{r plot-alluvial, fig.height=3.8, comment = F}

p_alluvial

```

What are the species prevailing in regeneration?


```{r plot-alluvial2, fig.height=7, comment = F}

p_alluvial2

```
